---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
library(dplyr)
datax <- data %>%
     group_by(date) %>%
     summarise_at(vars(-interval), list(sum = sum))
hist(datax$sum,xlab = "Total number of steps per day", ylab = "Number of days", main = "Total number of steps taken each day")
meanx <- mean(datax$sum, na.rm = TRUE)
medianx <- median(datax$sum, na.rm = TRUE)
print(paste("The mean is", meanx), quote = FALSE)
print(paste("The median is", medianx), quote = FALSE)
```

## What is the average daily activity pattern?

```{r}
datay <- data %>%
      group_by(interval) %>%
      summarise_at(vars(-date), list( mean = mean), na.rm = TRUE)
with(datay, plot(interval,mean, type = "l", ylab = "Steps Mean", xlab = "Interval", main = " Average Daily Activity"))
```

## Imputing missing values
```{r}
nacount <- sum(is.na(data$steps))
print(paste("There are",nacount, "NA values"), quote = FALSE)

```
I chose to replace the NA values with the mean of the steps column (Every day,
Every interval)

```{r}
data2 <- data
data2[is.na(data2[,1]),1]<-mean(data2[,1], na.rm=TRUE)
dataw <- data2 %>%
     group_by(date) %>%
     summarise_at(vars(-interval), list(sum = sum))
hist(dataw$sum,xlab = "Total number of steps per day", ylab = "Number of days", main = "Total number of steps taken each day")
meanw <- mean(dataw$sum, na.rm = TRUE)
medianw <- median(dataw$sum, na.rm = TRUE)
print(paste("The mean is", meanw), quote = FALSE)
print(paste("The median is", medianw), quote = FALSE)
```

The impact of the change can be seen on the histogram, the frequency of total number
of days between 10.000 and 15.000 has increased greatly. Also, the median and the
mean is now the same, due to the change.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
data$date<-as.Date(data$date)
data3 <- data %>%
        mutate(Day = ifelse(weekdays(data$date)=="sabado" |            weekdays(data$date)=="domingo", "Weekend", "Weekday"))
datav <- data3 %>%
      group_by(interval, Day) %>%
      summarise_at(vars(-date), list( mean = mean), na.rm = TRUE)
with(datav, xyplot(mean ~ interval | Day, type = "l", main = "Total Number of Steps within Intervals by dayType",xlab = "Daily Intervals", ylab = "Average Number of Steps"))
```

